---
title: "Week 7 Piping"
author: "WCW"
date: "2026-02-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#load the tidyverse package

requiredPackages = c('tidyverse', 'nlme')
for(p in requiredPackages){
  if(!require(p,character.only = TRUE)) install.packages(p)
  library(p,character.only = TRUE)
}
```

## Piping in tidyverse: back to basics (again)

A pipe looks like this: %>% 
And can be used like this:

```{r cars}
iris2<- iris %>% 
  filter(iris$Species == 'setosa')
```

## .... What?

piping combines functions into a sequential pipeline, each step is performed step by step

So the above says 
    1. Make a new dataframe called 'iris2' that is a copy of iris
    2. After making this dataframe SUBSET it to just 'setosa' species.
    
    
## How is this different from what we just learned like a week ago?

Well you could do...

```{r pressure, echo=FALSE}
iris3 <- iris[iris$Species == 'setosa',]

```

## ... Right.

But what if we want to do more 
```{r p1, echo=FALSE}
iris3 <- iris[iris$Species == "setosa" & iris$Sepal.Length > 5, ]
rm(list = ls())
```
## ... ORRRRR

```{r p2, echo=FALSE}
iris4 <- iris[iris$Species == "setosa" & iris$Sepal.Length > 5 & iris$Petal.Width>0.3 & iris$Sepal.Width>4, ]

```

## Wow that sucks to look at, what does it even say?


I know right? 


That's why the tidyverse exists

```{r p3, echo=FALSE}


iris2<- iris %>% 
  filter(Species == 'setosa') %>% 
  filter(Sepal.Length > 5) %>% 
  filter(Petal.Width > 0.3) %>% 
  filter(Sepal.Width > 4)

```

These two code chunks do the exact same thing, but one is easier for you to read 


## Okay sure it's a bit cleaner... But why did you write filter so many times

Because it's an example, silly.

There are plenty of functions to use in the tidyverse.

Check out this cheat sheet https://github.com/rstudio/cheatsheets/blob/main/data-transformation.pdf


## Lets look at them 
```{r p4, echo=FALSE}



iris2<- iris %>% 
  filter(Species == 'setosa') %>% 
  mutate(Petal.area = Petal.Length*Petal.Width/4 * pi^2 ) 
# NOTE THIS IS NOT HOW YOU GET SEPAL AREA

```

What this says is:
    1. Make a new dataframe called 'iris2' that is a copy of iris
    2. After making this dataframe SUBSET it to just 'setosa' species.
    3. THEN CREATE a and compute a new column 
    
In practice you're more likely to write something like this:
    
```{r p5, echo=FALSE}



iris2<- iris %>% 
  select(Petal.Length,Petal.Width,Species) %>% 
  mutate(Petal.area = Petal.Length*Petal.Width/4 * pi^2 ) %>% 
  group_by(Species) %>% 
  summarize(avg = mean(Petal.area), st_dev = sd(Petal.area))

```

So now we 
    1. Make a new dataframe
    2. THEN CREATE and compute the area
    3. THEN we group by species
    4. THEN we find the mean and standard deviation for Petal area
    
    
YOUR TURN :)

We're using spruce data again, first let's look in it

```{r homeworkbegin, echo = F}
rm(list = ls())
head(Spruce)

spruce<- Spruce
```
Okay 4 columns AND its already grouped


### QUESTION 1: Choose the functions that will finish the code below. This code chunk will make a new spruce dataframe that only includes N. REPLACE THE ______ with the function..

Hint: You need to CREATE a new function and then SUBSET it

```{r homework1}
spruce <- Spruce %>% 
  ungroup() %>% 
  _______(treatment = substr(Tree, 0,1)) %>% 
  _______(treatment == 'N')


```

### QUESTION 2: Summarize the original dataframe to plot and days get mean and standard deviation
```{r homework2}
sums<- Spruce %>% 
  group_by(plot,days) %>% 
  summarize(avg = mean(logSize), stdev = sd(logSize) )


```

### Question 3: Plot the average logSize for each plot as a boxplot and make each plot a different color use the new spruce dataframe I make you

There are many ways to do this with ggplot()
```{r homework3}
spruce<-Spruce %>% 
  mutate(time =if_else(days < 366, "year 1", "year 2")) %>% 
  mutate(unlog = exp(logSize))
ggplot(spruce, aes(x = plot,y = unlog, color = plot))+
  geom_boxplot()+
  facet_wrap(~time)

```


##Final Question: If treatment = N meant Nitrogen was added to the soil, and your sums dataframe shows that plots 3 and 4 received N treatment... then what is the previous graph showing in terms of nitrogen fertilization. Is there a difference between year 1 and year 2?


Answer:








#knit this into an html and submit to Canvas